Predicting I/O Performance in HPC Using Artificial Neural Networks

### 2016-08-22 Paper/paper.tex
kleine Anpassungen

		The prediction of file access times is an important part for the modeling of supercomputer's storage systems. 
		These models can be used to develop analysis tools which support the users to integrate efficient I/O behavior.\\
		In this paper, we analyze and predict the access times of a Lustre file system from the client perspective.
		Therefore, we measure file access times in various test series and developed different models for predicting access times. 
		The evaluation shows that models utilizing artificial neural networks achieved about 30\% smaller average prediction errors than linear models. 
		A phenomenon in the distribution of file access times is of particular interest:
		File accesses with identical parameters show several typical access times.\\
		The typical access times usually differ by orders of magnitude and can be explained with a different processing of the file accesses in the storage system -- an alternative I/O path.\\
		We investigate a method to automatically determine the alternative I/O path and quantify the significance of knowledge about the internal processing.
		It is shown that the prediction error is improved significantly with this approach.

@@

	\subsection{Model of the I/O path}
	The internal processing of a file access in the storage system can be viewed using the I/O path.	
	The I/O path of a file access is the interaction between all components that are involved in the transfer of data read or written for its execution.
	For example, for a node local file system this includes operating system, memory and storage device / medium. Remote systems add network	traffic and usually server sided components.
	Usually, in parallel file systems the I/O of the client is transferred through operating system, client side file system modules, network, file system servers, storage devices and ultimately media.
	In several of these steps, data may be cached in memory and optimizations can take place.

@@

	\begin{figure}[t]
		\centering
		\begin{minipage}[b]{0.47\textwidth}
			\includegraphics[width=\textwidth]{src/periodic_peaks.png}
			\caption{Small sequence of read accesses with periodic peaks in access time}
			\label{periodic}
		\end{minipage}
		\hfill
		\begin{minipage}[b]{0.47\textwidth}
			\includegraphics[width=\textwidth]{src/sporadic_peaks.png}
			\caption{Small sequence of read accesses with sporadic peaks in access time}
			\label{sporadic}
		\end{minipage}
	\end{figure}

	With knowledge about the periods of peaks a model could make better access time predictions.
	The ema-model is supposed to use its input information of previous data throughput for that.
	The data throughput has peaks the same period as the access time. We use the data throughput instead of the access times of measurements, because an analysis of the data throughput could find periodic behavior of system performance even in a sequence with different access sizes. 
	Internally the model calculates the exponential moving average (EMA) of data throughput of measurement $i$ as:
	\begin{equation}
	EMA(i) = 0.5 \cdot t(i)+ 0.5 \cdot EMA(i-1)
	\end{equation}
	With $t(i)$ the access time of measurement $i$.
	The influence of data throughput of a measurement is exponentially decreasing in the series of EMAs.
	In Figure \ref{periodic} the EMA-function for a function with periodic peaks can be seen.
	\begin{figure}[t]
		\centering
		\begin{minipage}[b]{0.47\textwidth}
			\includegraphics[width=\textwidth]{src/ema.png}
			\caption{Exponential moving average (in red) for a function with periodic peaks}
			\label{periodic}
		\end{minipage}
	\end{figure}	
	The ema-model can use $EMA(n-1)$ for the prediction of access time of measurement $n$.
	If the model is able to find threshold values of the EMA-function after which another peaks follows, it can use this input information for better access time predictions on sequences with periodic peaks in access time. Depending on how relevant periodic behavior of access times is in our measurements, this additional input information could improve the model.\medskip

	Additional models that were examined in \cite{VVEIHUDVVN15}, but won't be analyzed in further detail in this paper, are summarized in the following list:
	\begin{itemize}
		\item Using multiple linear regression, we analyzed the performance of more baseline models. We used access size and delta-offset as tuple and also access size, delta-offset and access type as triple to find a mapping to access time.
		Despite having more information to work with both models achieve equal or worse deviations from the measured access times than the simple linear regression which only considers access size.
		If access type and delta-offset have a meaningful correlation to the access time of a file access, it requires more sophisticated tools than linear regression for exploitation.
		\item An ANN-model receiving knowledge about the previous measurement was examined as well.
		Additionally to the input information of the simple ANN-model the access size, access type, delta-offset and also access time of the previous file access was available for this model.
		Conceptually this model might be able to compare the actual performance on the previous file access with its own prediction for it. If the actual performance deviates from the predicted performance this model could exploit this knowledge to adapt its prediction for the following file access. In case of a prolonged period of high workload this adaptation might have positive affect on the predictions.
		The results of this model, however, were throughout worse compared to the simple ANN-model which has less information to work with, even with quite complex network structures available. 
		Therefore we conclude that the performance of the direct predecessor of a file access can't or is to complicated to be exploited for an improved prediction.
		\item An ANN-model similar to the error-class-model described above using error classes obtained from the residues of the simple ANN-model instead of residues from the linear regression was examined as well. This approach lead to very similar prediction performance to the model using error classes from linear regression. On the data set of sequential file accesses this model achieved with $2.4\cdot10^{-5}$ to $2.0\cdot10^{-5}$ a slightly worse and on measurements with random access pattern with  $8.9\cdot10^{-4}$ to $10.3\cdot10^{-4}$ a slightly better mean average error.
	\end{itemize}

@@

	As expected, sequential file accesses are on average much faster than random accesses especially for the reading case, also random reads are slower than writes. In Figure \ref{correlations}, we consider the correlations between file access parameters and the resulting access time, a strong correlation can be exploited for access time prediction.
	\tab{correlations}{Correlations between file access parameters and access time}{
		\centering
		\scriptsize
		\begin{tabular}{|r|r|r|}\hline%
			Attribute & Sequential access pattern & Random access pattern\\\hline\hline
			\csvreader[late after line=\\\hline]%
			{src/data_summary_correlations.csv}{Attribut=\Attribut,seq = \seq, rnd = \rnd}%
			{\Attribut & \seq &\rnd}%
		\end{tabular}
	}

@@

	\begin{figure}[b]
		\centering
		\begin{minipage}[b]{0.47\textwidth}
			\includegraphics[width=\textwidth]{src/plot_SizeSorted_log_read_seq.png}
			\caption{Measurements of sequential reads}
			\label{read_seq}
		\end{minipage}
		\hfill
		\begin{minipage}[b]{0.47\textwidth}
			\includegraphics[width=\textwidth]{src/plot_SizeSorted_log_write_seq.png}
			\caption{Measurements of sequential writes}
			\label{write_seq}
		\end{minipage}
	\end{figure}
	\begin{figure}[t]
		\centering
		\begin{minipage}[b]{0.47\textwidth}
			\includegraphics[width=\textwidth]{src/plot_SizeSorted_log_read_rnd.png}
			\caption{Measurements of random reads}
			\label{read_rnd}
		\end{minipage}
		\hfill
		\begin{minipage}[b]{0.47\textwidth}
			\includegraphics[width=\textwidth]{src/plot_SizeSorted_log_write_rnd.png}
			\caption{Measurements of random writes}
			\label{write_rnd}
		\end{minipage}
	\end{figure}

>>>>

		The prediction of file access times is an important part for the modeling of supercomputer's storage systems. 
		These models can be used to develop analysis tools which support the users to integrate efficient I/O behavior.\\
		In this paper, we analyze and predict the access times of a Lustre file system from the client perspective.
		Therefore, we measured file access times in various test series and developed different models for predicting access times. 
		The evaluation shows that models utilizing artificial neural networks achieved about 30\% smaller average prediction errors than linear models. 
		A phenomenon in the distribution of file access times is of particular interest:
		File accesses with identical parameters show several typical access times.\\
		The typical access times usually differ by orders of magnitude and can be explained with a different processing of the file accesses in the storage system -- an alternative I/O path.\\
		We investigate a method to automatically determine the alternative I/O path and quantify the significance of knowledge about the internal processing.
		It is shown that the prediction error is improved significantly with this approach.

@@

	\subsection{Model of the I/O path}
	The internal processing of a file access in the storage system can be viewed using the I/O path.	
	The I/O path of a file access is the interaction between all components that are involved in the transfer of data read or written for its execution.
	For example, for a node local file system this includes operating system, memory and storage device/medium. Remote systems add network	traffic and usually server sided components.
	Usually, in parallel file systems the I/O of the client is transferred through operating system, client side file system modules, network, file system servers, storage devices and ultimately media.
	In several of these steps, data may be cached in memory and optimizations can take place.

@@

	\begin{figure}[t]
		\centering
		\begin{minipage}[b]{0.47\textwidth}
			\includegraphics[width=\textwidth]{src/periodic_peaks.png}
			\caption{Small sequence of read accesses with periodic peaks in access time}
			\label{periodic}
		\end{minipage}
		\hfill
		\begin{minipage}[b]{0.47\textwidth}
			\includegraphics[width=\textwidth]{src/sporadic_peaks.png}
			\caption{Small sequence of read accesses with sporadic peaks in access time}
			\label{sporadic}
		\end{minipage}
	\end{figure}
	With knowledge about the periods of peaks a model could make better access time predictions.
	The ema-model is supposed to use its input information of previous data throughput for that.
	The data throughput has peaks the same period as the access time. We use the data throughput instead of the access times of measurements, because an analysis of the data throughput could find periodic behavior of system performance even in a sequence with different access sizes. 
	Internally the model calculates the exponential moving average (EMA) of data throughput of measurement $i$ as:
	\begin{equation}
	EMA(i) = 0.5 \cdot t(i)+ 0.5 \cdot EMA(i-1)
	\end{equation}
	With $t(i)$ the access time of measurement $i$.
	The influence of data throughput of a measurement is exponentially decreasing in the series of EMAs.
	In Figure \ref{periodic} the EMA-function for a function with periodic peaks can be seen.
	%\begin{figure}[t]
	%	\centering
	%	\begin{minipage}[b]{\textwidth}
	%		\includegraphics[width=0.47\textwidth]{src/ema.png}
	%		\caption{Exponential moving average (in red) for a function with periodic peaks}
	%		\label{periodic}
	%	\end{minipage}
	%\end{figure}
	\fig{width=.47\textwidth}{src/ema.png}{Exponential moving average (in red) for a function with periodic peaks}	
	The ema-model can use $EMA(n-1)$ for the prediction of access time of measurement $n$.
	If the model is able to find threshold values of the EMA-function after which another peaks follows, it can use this input information for better access time predictions on sequences with periodic peaks in access time. Depending on how relevant periodic behavior of access times is in our measurements, this additional input information could improve the model.
	\medskip

	Additional models that were examined in \cite{VVEIHUDVVN15}, but won't be analyzed in further detail in this paper, are summarized in the following list:
	\begin{itemize}
		\item Using multiple linear regression, we analyzed the performance of other baseline models. We used access size and delta-offset as tuple and also access size, delta-offset and access type as triple to find a mapping to access time.
		Despite having more information to work with both models achieve equal or worse deviations to the measured access times than the simple linear regression which only considers access size.
		If access type and delta-offset have a meaningful correlation to the access time of a file access, they require more sophisticated tools than linear regression to be exploited.
		\item An ANN-model receiving knowledge about the previous measurement was examined as well.
		Additionally to the input information of the simple ANN-model the access size, the access type, the delta-offset and also the access time of the previous file access was given as input for this model.
		Conceptually this model might be able to compare the actual performance on the previous file access with its own prediction for it. If the actual performance deviates from the predicted performance this model could exploit this knowledge to adapt its prediction for the following file access. In case of a prolonged period of high workload this adaptation might have positive affect on the predictions.
		The results of this model, however, were throughout worse compared to the simple ANN-model which has less information to work with, even though quite complex network structures were used in the learning process (16 hidden layers with 17 neurons each for the sequential access pattern and 14 hidden layers with 17 neurons each for the random access pattern achieved the best results).
		Therefore we conclude that the performance of the direct predecessor of a file access can't or is to complicated to be exploited for an improved prediction.
		\item An ANN-model similar to the error-class-model described above, using error classes obtained from the residues of the simple ANN-model instead of residues from the linear regression, was examined as well. This approach lead to very similar prediction performance to the model using error classes from linear regression. On the data set of sequential file accesses this model achieved with $2.4\cdot10^{-5}$ to $2.0\cdot10^{-5}$ a slightly worse and on measurements with random access pattern with  $8.9\cdot10^{-4}$ to $10.3\cdot10^{-4}$ a slightly better mean average error.
	\end{itemize}

@@

	As expected, sequential file accesses are on average much faster than random accesses especially for the reading case, also random reads are slower than writes. In Table \ref{correlations}, we consider the correlations between file access parameters and the resulting access time, a strong correlation can be exploited for access time prediction.
	\tab{correlations}{Correlations between file access parameters and access time}{
		\centering
		\scriptsize
		\begin{tabular}{|r|r|r|}\hline%
			Attribute & Sequential access pattern & Random access pattern\\\hline\hline
			\csvreader[late after line=\\\hline]%
			{src/data_summary_correlations.csv}{Attribut=\Attribut,seq = \seq, rnd = \rnd}%
			{\Attribut & \seq &\rnd}%
		\end{tabular}
	}

@@

	\begin{figure}[b]
		\centering
		\begin{minipage}{0.47\textwidth}
			\includegraphics[width=\textwidth]{src/plot_SizeSorted_log_read_seq.png}
			\caption{Measurements of sequential reads}
			\label{read_seq}
		\end{minipage}
		\hfill
		\begin{minipage}{0.47\textwidth}
			\includegraphics[width=\textwidth]{src/plot_SizeSorted_log_write_seq.png}
			\caption{Measurements of sequential writes}
			\label{write_seq}
		\end{minipage}
	\end{figure}
	\begin{figure}[b]
		\centering
		\begin{minipage}{0.47\textwidth}
			\includegraphics[width=\textwidth]{src/plot_SizeSorted_log_read_rnd.png}
			\caption{Measurements of random reads}
			\label{read_rnd}
		\end{minipage}
		\hfill
		\begin{minipage}{0.47\textwidth}
			\includegraphics[width=\textwidth]{src/plot_SizeSorted_log_write_rnd.png}
			\caption{Measurements of random writes}
			\label{write_rnd}
		\end{minipage}
	\end{figure}
